---
title: "take home exercise 4"
author: "Xu Lin"
date: "MAR 9, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  warning: false
editor: visual
---

# Purpose:

we want to find out the present rent price in Singapore Market.

#Data preparation：
The data we download from [SG HDB website](https://www.ura.gov.sg/property-market-information/pmiResidentialRentalSearch).
```{r}
pacman::p_load(sf, tmap, tidyverse, dplyr, readr, plotly, seriation, dendextend, heatmaply)
```

```{r}
Retentialdata <- read_csv("data/ResidentialRental/20241.csv")
```

we change the variable more easy to read.
```{r}
Retentialdata2024 <- Retentialdata %>%
  select(
    PostalDistrict = `Postal District`, 
    NoOfBedroom = `No of Bedroom`, 
    MonthlyRent = `Monthly Rent ($)`,
    StreetName  = `Street Name`,
    ProjectName1 = `Project Name`)
```

```{r}
postdistrictdata <- read_csv("data/ResidentialRental/realis2018.csv") %>%
  select(postalDistrict2 = "Postal District", 
         PlanningArea = "Planning Area",
         PlanningRegion = "Planning Region",
         ProjectName2 = "Project Name") %>%
  distinct()
```

#Data Merge
We use left_join to merge the data, because we want to add "PlanningArea".
```{r}
Retentialdata2024merged <- Retentialdata2024 %>%
  left_join(postdistrictdata, by = c("ProjectName1" = "ProjectName2")) %>%
  drop_na() %>%
  mutate(PlanningArea = case_when(
    PlanningArea == "YISHUN" ~ "Yishun",
    TRUE ~ PlanningArea 
  )) 
```

We want to select the NoOfBedroom 1-4.
```{r}
filtered_data <- Retentialdata2024merged %>%
  filter(NoOfBedroom %in% 1:4) %>%
  group_by(PlanningArea, NoOfBedroom) %>%
  summarise(Median_Rent = median(MonthlyRent, na.rm = TRUE)) %>%
  ungroup()  %>%
  pivot_wider(names_from = NoOfBedroom, values_from = Median_Rent)
```

```{r}
rent2024 <- filtered_data %>%
  pivot_longer(-PlanningArea, names_to = "NoOfBedroom", values_to = "Median_Rent")
```

```{r}
rent2024room1 <- rent2024 %>% filter(NoOfBedroom == 1)
```

The LLNEW downloads from Tableau, then we change the format that R can read. 
```{r}
LLNEW <- read_csv("data/ResidentialRental/LLNEW.csv")%>%
  select(PlanningArea1 = "Planning Area",
       Latitude = "Latitude (generated)",
       Longitude = "Longitude (generated)")
```

```{r}
rent2024room1_joined <- rent2024room1 %>%
  left_join(LLNEW, by = c("PlanningArea" = "PlanningArea1"))
```

#MAP
We use the sf package to convert the data to geospatial data. 
```{r}
rent2024room1_joined_sf <- st_as_sf(rent2024room1_joined,
                      coords = c("Longitude",
                      "Latitude"),
                      crs = 4326) %>%
  st_transform(crs = 3414)
```

We use the mpsz dataset in the mapping function to visualize geographic data.
```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL") %>%
  st_transform(crs = 3414)
```

```{r}
p1<-tmap_options(check.and.fix = TRUE)
   tmap_mode("view")
   tm_shape(mpsz) +
     tm_borders() +
   tm_shape(rent2024room1_joined_sf) +
     tm_dots(col= "Median_Rent")
   tmap_mode("plot")
```

```{r}
p2<-ggplot(rent2024, aes(x = PlanningArea, y = NoOfBedroom, fill = Median_Rent)) +
  geom_tile() + 
  labs(x = "Planning Area", y = "Number of Bedrooms", fill = "Median Monthly Rent") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
plotly_p <- ggplotly(p2, tooltip = "fill")
plotly_p
```





