---
title: "take home exercise 3"
author: "Xu Lin"
date: "Feb 11, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
editor: visual
---

# Background 
 As we know Singapore is situated near the equator and it has a typically tropical climate. It has 2 two monsoon seasons, one is Northeast Monsoon(December to early March); the other is Southwest Monsoon(une to September). In this task, we will choose December to analysis the temperature variation.

# Purpose 
Daily mean temperatures are projected to increase by 1.4 to 4.6, while annual mean temperatures rose at an average rate of 0.25 per decade.
We want to find out the reasons. 

```{r}
pacman::p_load(tidyverse)
```

```{r}
Data_1983_12TEM <- read_csv ("data/DAILYDATA_S24_198312.csv")
Data_1993_12TEM <- read_csv("data/DAILYDATA_S24_199312.csv") 
Data_2003_12TEM <- read_csv("DAILYDATA_S24_200312.csv") 
Data_2013_12TEM <- read_csv("DAILYDATA_S24_201312.csv") 
Data_2023_12TEM <- read_csv("DAILYDATA_S24_202312.csv") 
```





































