---
title: "take home exercise 3"
author: "Xu Lin"
date: "Feb 11, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
editor: visual
---

# Background

As we know Singapore is situated near the equator and it has a typically tropical climate. It has 2 two monsoon seasons, one is Northeast Monsoon(December to early March); the other is Southwest Monsoon(une to September). In this task, we will choose December to analysis the temperature variation.

# Purpose

Daily mean temperatures are projected to increase by 1.4°C to 4.6°C, while annual mean temperatures rose at an average rate of 0.25°C per decade. We want to find out the reasons.

```{r}
pacman::p_load(tidyverse, ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse, ggdist, ggridges, colorspace,gridExtra)
```

# Import data

I choose the Changi station, December 1983, 1993, 2003, 2013, 2023' temperature data by grouping the data manually. R can not read the degree, so I moved it out.

```{r}
TEMData <- read_csv("data/DEC1983-2023TEM.csv") 
```

```{r}
glimpse(TEMData)
```

# Data description &  Process of thinking.

Ridgeline plot(also known as Joyplot) is a method to display the distribution of a numeric value for several groups. We use this method to display the December of 1983, 1993, 2003, 2013, 2023's daily mean temperature.

```{r}
ggplot(data = TEMData, 
       aes(x = `Mean Temperature`, y = as.factor(Year), 
           fill = after_stat(x))) + 
    geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp (°C)",
                       option = "C") +
  scale_x_continuous(
    name = "Mean Temperature",
    expand = c(0, 0)
  ) +
  scale_y_discrete(
    expand = expansion(add = c(0.2, 2.6))
  ) +
  theme_ridges() +
  labs(title = "Dec Temperature Distribution by Year",
    y = "Year")
```

```{r}
ggplot(TEMData,
       aes(x = `Mean Temperature`, y = as.factor(Year), 
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability",
                       direction = -1) +
  theme_ridges()+
  labs(title = "Dec Temperature Distribution by Year",
    y = "Year")
```

```{r}
ggplot(TEMData,
       aes(x = `Mean Temperature`, y = as.factor(Year),
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = 4,
    quantile_lines = TRUE) +
  scale_fill_viridis_d(name = "Quartiles") +
  theme_ridges() +
  labs(title = "Dec Temperature Distribution by Year",
    y = "Year")
```

```{r}
ggplot(TEMData,
       aes(x = `Mean Temperature`, y = as.factor(Year), 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
    ) +
  scale_fill_manual(
    name = "Probability",
    values = c("#FF0000A0", "#A0A0A0A0", "#0000FFA0"),
    labels = c("(0, 0.025]", "(0.025, 0.975]", "(0.975, 1]")
  ) +
  theme_ridges()+
  labs(title = "Dec Temperature Distribution by Year",
    y = "Year")
```

```{r}
ggplot(TEMData,
       aes(x = as.factor(Year), y = `Mean Temperature`)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  labs(title = "Dec Temperature Distribution by Year",
    x = "Year")
```

```{r}
ggplot(TEMData,
       aes(x = as.factor(Year), y = `Mean Temperature`)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  labs(title = "Dec Temperature Distribution by Year",
    x = "Year")
```

```{r}
TEMData_means <- TEMData %>%
  group_by(Year) %>%
  summarise(MeanTemp = mean(`Mean Temperature`, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
ggplot(TEMData,
       aes(x = as.factor(Year), y = `Mean Temperature`)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 0.1)+
  labs(title = "Dec Temperature Distribution by Year",
    x = "Year") +
  geom_text(data = TEMData_means, aes(label = round(MeanTemp, 1), y = MeanTemp), 
          nudge_x = 0.25, hjust = 0, check_overlap = TRUE)
```
# Final result
```{r}
TEMData_max <- aggregate(`Mean Temperature` ~ Year, data = TEMData, max)
TEMData_min <- aggregate(`Mean Temperature` ~ Year, data = TEMData, min)
```

```{r}
p1 <-ggplot(TEMData, aes(x = as.factor(Year), y = `Mean Temperature`)) +
  labs(y = "Mean Temperature (°C)", x = "Year") +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 0.1) +
  coord_flip() +
  theme_economist() +
  labs(title = "Dec Temperature Distribution by Year",
       x = "Year") +
  geom_text(data = TEMData_means, aes(label = round(MeanTemp, 1), y = MeanTemp), 
            nudge_x = 0.25, hjust = 0, check_overlap = TRUE) +
  geom_text(data = TEMData_max, aes(label = `Mean Temperature`, y = `Mean Temperature`),
            nudge_x = 0.25,nudge_y = -0.25, hjust = 0, check_overlap = TRUE) +  
  geom_text(data = TEMData_min, aes(label = `Mean Temperature`, y = `Mean Temperature`),
            nudge_x = 0.25, hjust = 0, check_overlap = TRUE)    
```

```{r}
TEMData_diff <- data.frame(
  Year = TEMData_max$Year,
  Diff = TEMData_max$`Mean Temperature` - TEMData_min$`Mean Temperature`
)
```

```{r}
p2 <- ggplot(TEMData_diff, aes(x = as.factor(Year), y = Diff)) +
  geom_bar(stat = "identity", fill = "grey")+
  geom_text(aes(label = Diff), position = position_dodge(width = 0.9), hjust = 0.8) +
  coord_flip() +
  theme_economist() +
  labs(x = "", y = "Tem Diff (°C)")
```

```{r}
combined_plot <- p1 + p2 + plot_layout(widths =c(3, 1))
combined_plot
```

# Summary 
For this purpose, we want to find out the daily mean temperatures increase by 1.4°C to 4.6°C, while annual mean temperatures rose at an average rate of 0.25°C per decade.

In the chart, you can see the minimal daily mean temperature increase from 25.4°C to 27.3°C from December 1983 to 2023. The trend of minimal daily mean temperature increase from 23°C to 26°C, while the maximal daily mean temperature from 27°C to 28.4°C from the left chart. From the right chart, you can see the daily mean temperature narrow form 4°C to 2.4°C. So we can conclude that the temperature is rising, and the daily temperature range is decreasing, with the rate of change accelerating.




















